---
title: "Data Set Up"
author: "Jackson Sorensen"
date: "6/6/2019"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Reading in data tables
This code is meant to help set up the data structures that will be used in later analyses and save them as R data structures to make for more efficient running later. 
## Microcosm metadata map

Reading in the metadata map for the microcosms

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
setwd("~/GitHub_Repos/ShadeLab/Sorensen_Microcosm/R")
map <- read_csv("InputFiles/Microcosm_Map.csv")
map$Date <- as.Date(map$Date, "%m/%d/%y")
```

## Reading in OTU table and collapsing
When we performed sequencing, we submitted 308 samples that were pooled together and then sequenced on 4 different MiSeq Runs. The following code reads in the otu table that has 4 columns for every sample submitted and collapses these 4 columns down to a single column for each sample by summing the 4. 
```{r,message=FALSE}
AllSamples <- read_table("InputFiles/AllSamples.txt", col_names = FALSE)

map <- map[order(map$Sample),]

AllSamples$X2 <- rep(map$Sample, 4)

otu <- read_tsv("InputFiles/table_combined_merged_otus.txt", col_names = TRUE)

otu <- otu %>%
    gather(key = SequenceName, value = value, 2:ncol(otu)) %>% 
    spread_(key = names(otu)[1],value = 'value')



colnames(AllSamples) <- c("SequenceName", "Sample")

otu <- left_join(otu, AllSamples, by="SequenceName")
otu<- otu[,-1]

otu_collapsed <- otu %>% 
  group_by(Sample) %>%
  summarise_all(funs(sum))

otu_collapsed <- otu_collapsed %>%
    gather(key = OTU, value = value, 2:ncol(otu_collapsed)) %>% 
    spread_(key = names(otu_collapsed)[1],value = 'value')
otu_collapsed <- as.data.frame(otu_collapsed)
row.names(otu_collapsed) <- otu_collapsed$OTU
otu_collapsed <- otu_collapsed[,-1]

saveRDS(otu_collapsed, file = "DataStructures/otu_collapsed.RDS")
saveRDS(map, file="DataStructures/map.RDS")


```